### 2018-10-13 manuscript.rmd
added file stubs for R analyses

---
csl: plos.csl
output:
  github_document: default
  html_document: default
  pdf_document: default
bibliography: manuscript.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=TRUE
	)
```

```{r preliminaries}
library( tidyverse )
library( magrittr )
```

# Rooting the animal tree of life

## Introduction

There has long been uncertainty about the relationship of Ctenophora to other animals [@Wallberg:2004ws].

## Published analyses

## References