### 2020-07-16 manuscript/explorations/phylobayes_explorations.rmd
some explorations of equilibrium frequency heterogeneity

---
output:
  html_document: default
  pdf_document: default
  word_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dpi = 300,
  cache = FALSE
  )
```

```{r preliminaries}
  library( tidyverse )
  library( magrittr )

  source( "../phylobayes.R" )

  # A colorblind friendly palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette:
  cbPalette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

  result_colors = c( cbPalette[2], cbPalette[4], cbPalette[1] )
  names( result_colors ) = c( "Ctenophora-sister", "Porifera-sister", "Unresolved" )

```

An exploration of phylobayes chain samples.

```{r}

chain = parse_phylobayes_chain("../tests/testthat/example.chain")

# Take just the last sample for now
pb = chain[[length(chain)]]

allocation_counts = table(pb@allocation)

# Building on https://www.statmethods.net/advstats/mds.html
d = dist( pb@frequencies )
fit = cmdscale(d,eig=TRUE, k=2)

categories = tibble(
  index = allocation_counts %>% names() %>% as.numeric(),
  count = c(allocation_counts),
  x = fit$points[,1],
  y = fit$points[,2]
)

categories %>% 
  ggplot() +
  geom_point(aes(x=x, y=y, size=count) ) +
  ggtitle("MDS plot of rate categories") +
  theme_classic()

```