### 2020-07-19 manuscript/explorations/phylobayes_explorations.rmd
parses unconstrained CAT chains

---
output:
  html_document: default
  pdf_document: default
  word_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = FALSE,
  dpi = 300,
  cache = FALSE
  )
```

```{r preliminaries}
  library( tidyverse )
  library( magrittr )
  library( gridExtra )

  source( "../phylobayes.R" )

  # A colorblind friendly palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette:
  cbPalette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

  result_colors = c( cbPalette[2], cbPalette[4], cbPalette[1] )
  names( result_colors ) = c( "Ctenophora-sister", "Porifera-sister", "Unresolved" )

```

An exploration of phylobayes chain samples.

```{r}

plot_chain = function( pb ){

  allocation_counts = table(pb@allocation)

  # Building on https://www.statmethods.net/advstats/mds.html
  d = dist( pb@frequencies )
  fit = cmdscale(d,eig=TRUE, k=2)

  categories = tibble(
    index = allocation_counts %>% names() %>% as.numeric(),
    count = c(allocation_counts),
    x = fit$points[,1],
    y = fit$points[,2]
  )

  gg_mds = 
  categories %>% 
    ggplot() +
    geom_point(aes(x=x, y=y, size=count) ) +
    ggtitle("MDS plot of rate categories") +
    theme_classic()

  gg_rank = 
    categories %>%
    arrange(desc(count)) %>%
    mutate(rank = 1:nrow(categories)) %>%
    ggplot() +
      geom_point(aes(x=rank, y=count) ) +
      ggtitle("Ordered count") +
      theme_classic()

  p = grid.arrange(
  	gg_rank,
  	gg_mds,
  	ncol = 2
  )

  return(p)
}

```

```{r}

chain = parse_phylobayes_chain( "../tests/testthat/example.chain" )

# Take just the last sample for now
pb = chain[[length(chain)]]

plot_chain( pb )

```

**Figure n60.** Plots for nCAT=60.

```{r}

chain = parse_phylobayes_chain( "Whelan2017_strict.phy_Poisson_CAT100_Chain_1_last500.chain" )

# Take just the last sample for now
pb = chain[[length(chain)]]

plot_chain( pb )

```

**Figure n100.** Plots for nCAT=100.

```{r}

chain = parse_phylobayes_chain( "Whelan2017_strict.phy_Chain1_last50000.chain" )

# Take just the last sample for now
pb = chain[[length(chain)]]

plot_chain( pb )

```

**Figure CAT.** Plots for CAT.